---
title: "Statistics 652 - Final"
author: "Ken Vu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_notebook: default
---


# Final

For the *Ozone* data from the R package *mlbench* try the following machine learning prediction algorithm that is useful for feature selection.

Read the paper [Feature Selection with the Boruta Package](https://www.jstatsoft.org/article/view/v036i11) and implement the algorithm.

Which features are most important as determined by the Boruta RandomForest Algorithm?

**Answers**:
Based on the results of the Boruta RandomForest algorithm, we find that the most important features are V1, V4, V5, V6, V7, V8, V9, V10, V11, V12, and V13.

# Code and Comments (Final):
Get all of the libraries first
```{r}
library(mlbench)
library(Boruta)
set.seed(300)
```

## 1) Load the data set
```{r}
data(Ozone)
data_raw <- Ozone
```

## 2) Explore the data set
```{r}
summary(data_raw)
```

### Investigate the missing values
```{r}
missing_values <- as.data.frame(sapply(data_raw, function(x) sum(is.na(x))))
colnames(missing_values) <- c("Number of NAs")
missing_values$freq <- round(missing_values[,c("Number of NAs")]/length(data_raw[,1]),3)

# Get columns where 10% of data r less is NAs
colnames_lessmiss <- rownames(missing_values[missing_values$freq <= 0.1,])
missing_values[]
```

### See if you can mean impute for some of the variables
```{r}
# Mean impute for the columns with 10% or less data missing
NA2mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
data_clean <- data_raw
data_clean[,colnames_lessmiss] <- lapply(data_clean[,colnames_lessmiss], NA2mean)

# Check the proportion of missing data to total row count again
missing_values_imputed <- as.data.frame(sapply(data_clean, function(x) sum(is.na(x))))
colnames(missing_values_imputed) <- c("Number of NAs")
missing_values_imputed$freq <- round(missing_values_imputed[,c("Number of NAs")]/length(data_clean[,1]),3)
missing_values_imputed
```

### Now, clean out the data of NAs
```{r}
data_clean <- na.omit(data_clean)
```

## 3) Train Boruta onto the data set.

We limit max runs to 12 in case there are attributes with importance too close to MSZA that might make algorithm indecisive about which attributes to consider important.
```{r}
Boruta.Short <- Boruta(V4 ~ ., data = data_clean, doTrace = 2, ntree = 500, maxRuns=12)
Boruta.Short
```

## 4) Now, try to do a tentative fix on the algorithm to sort out the tentative variables.
```{r}
Boruta.TentFix<- TentativeRoughFix(Boruta.Short)
Boruta.TentFix
```

## 5) Plot the results of Boruta
```{r}
plot(Boruta.TentFix, las=2)
```

